### 2019-11-26 paper.rmd
re-write intro

Person-oriented analysis involves a statistical approach to exploring
psychological constructs in a way that captures and simplifies complexity. This
approach can be used to consider the way in which psychological constructs, for
example, are experienced together and at once. Thus, conceptually, this approach
emphasizes how individuals go about interacting, socializing, and behaving in
this way, experiencing these activities as a whole person. Person-oriented
analysis, developed within developmental science, focuses on how individuals
have experiences in a holistic way and also how these ways can be identified
through common groups, or profiles.(Bergman & El-Khouri, 1997; Magnusson &
Cairns, 1996). One way to distinguish the approach is to consider it in contrast
to variable-centered approaches, such as those for which the covariability
between variables, rather than the relations between profiles of individuals, is
the focus (Bergman & Trost, 2006).

Though studies examining learning from a person-oriented perspective are not
very common, some examples include studies of intrinsic and extrinsic motivation
(Corpus & Wormington, 2014; Hayenga & Corpus, 2010), profiles of achievement
goals (see Wormington & Linnenbrink-Garcia, advance online publication, for a
review), and epistemic cognition (Trevors, Kendeou, Braten, & Braasch, 2017).
Recently, too, there have been excellent reviews of person-oriented analyses in
motivation and educational psychology (Linnenbrink-Garcia & Wormington, 2017).

In this article, our goal is to review person-oriented analysis focused on
getting started (on both knowing about and doing person-oriented analyses) in an
informed way. More specifically, we start with an overview of the methodology,
describe specific software tools, explain analytic choices made in the course of
their use, and conclude with general recommendations One challenge facing those
trying to learn about and apply person-oriented analysis in their research is
the multitude of terms used for similar analytic approaches or ideas. Throughout
this manuscript, we aim to point these out and explain when and why a variety of
terms have been used. Thus, this manuscript is written for an informed and
curious reader at any level interested in understanding this particular
approach.

>>>>

For teachers, psychologists, researchers, and friends and family member, people 
are highly-complex and not easily able to be characterized by a single 
characteristic or personality trait. In the social sciences, broadly, and in 
psychology, particularly, a statistical method that can be used to describe how 
people, in their individual particularities, may have similarities on the basis 
of some set of measures through which they can be grouped in meaningful, 
distinctive ways. This approach, which has a provenance in developmental 
approaches (Bergman & El-Khouri, 1997; Magnusson & Cairns, 1996), though is now 
widely used in educational, social, personality, and other domains of psychology, 
is an example of a general mixture model (Harring & Hodis, 2016; Pastor, Barron, 
Miller, & Davis, 2007).

In this tutorial, we aim to describe one of the most commonly-used--and relevant 
to psychologists--application of the general mixture model, to cases for which 
all of the variables for which (relatively) homogeneous groups are identified 
from among a (relatively) heterogeneous sample are continuos, *latent profile 
analysis* (LPA). After describing the method and some examples of its use, we 
provide a tutorial for carrying out LPA in the context of a freely-available, 
open-source statistical software package we created for R (R Core Team, 2019), 
tidyLPA. Finally, we offer some informed recommendations for researchers aiming
to use LPA in their applied work, and conclude with reflections on the role of 
statistical software--especially software that is freely-available, open-source, 
and highly-performant--in the psychological sciences.

<!-- Though studies examining learning from a person-oriented perspective are not -->
<!-- very common, some examples include studies of intrinsic and extrinsic motivation -->
<!-- (Corpus & Wormington, 2014; Hayenga & Corpus, 2010), profiles of achievement -->
<!-- goals (see Wormington & Linnenbrink-Garcia, advance online publication, for a -->
<!-- review), and epistemic cognition (Trevors, Kendeou, Braten, & Braasch, 2017). -->
<!-- Recently, too, there have been excellent reviews of person-oriented analyses in -->
<!-- motivation and educational psychology (Linnenbrink-Garcia & Wormington, 2017). -->
